---
title: "ADSC2020 Logistic Regression Examples"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Example 1 from slides

survived: no or yes.\
sex: female or male\
age: in years (and for some children, fractions of a year); age is missing for 263 of the passengers. passengerClass: 1st, 2nd, or 3rd class.

```{r}


library(carData)
data("TitanicSurvival")


TitanicS <- na.omit(TitanicSurvival)

# LogitModel <- glm(binary.response ~ Var.1 + Var.2 + ...
# + Var.j, family = binomial(link = "logit"), data = data)


```

## Example 2 from slides

```{r}

set.seed(2020)

TitanicSurvival$X1 <- rnorm(nrow(TitanicSurvival),2,2)
#relevel(passengerClass,"2nd")

```

### Linearity I

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# Model values #
probabilities <- predict(LogitModel, type = "response")

# Numeric variables #
mydata <- data %>%
na.omit() %>%
dplyr::select_if(is.numeric)
predictors <- colnames(mydata)

# transformed relationship #
mydata <- mydata %>%
mutate(logit = log(probabilities/(1-probabilities))) %>%
gather(key = "predictors", value = "predictor.value",-logit)

# generate the plots #

ggplot(mydata, aes(logit, predictor.value))+
geom_point(size = 0.5, alpha = 0.5) +
geom_smooth(method = "loess") +
theme_bw() +
facet_wrap( predictors, scales = "free_y")

```

### Linearity II

```{r}

TitanicSurvival %>%
  na.omit() %>% 
mutate(comp_res = coef(LogitModel)["variable.name"]*variable.name +
  residuals(LogitModel, type = "working")) %>%
ggplot(aes(x = variable.name, y = comp_res)) +
geom_point() +
geom_smooth(color = "red", method = "lm", linetype = 2,
se = F) +
geom_smooth(se = F)




```

## Example 3 from slides

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)



```

## Example 4 from slides

```{r}

library(car)
# vif()
```

## Example 5 from slides

```{r}

# plot(LogitModel,3)
# plot(LogitModel,4)

# cooks.distance(LogitModel)




```

## Example 6 from slides

```{r}

# deviance()
# anova(model.simple, model.complex, test='LR')
# BIC()
# anova(lm2,lm1,test='LR')

```

### Stepwise Selection

```{r}

intercept.model <- glm(survived ~ 1, data = na.omit(TitanicSurvival), family = binomial)
full.model <- glm(survived ~., data = na.omit(TitanicSurvival), family = binomial)

# Does not need to be all variables, can be a set under consideration.
both <- step(intercept.model, direction="both", scope=formula(full.model), trace=0)
# trace = 1 Shows each step
both$anova #Shows the results
both$coefficients #Shows the estimates


# BIC #

bothBIC <- step(intercept.model, direction="both", scope=formula(full.model), trace=0, k = log(nrow(na.omit(TitanicSurvival)))) 
# trace = 1 Shows each step
bothBIC$anova #Shows the results

bothBIC$coefficients #Shows the estimates




```

## Example 7 from slides

```{r}
set.seed(2020)

library(caret)

train_model <- trainControl(method = "cv", number = 10)

# Train the model
modelT <- train(
  survived ~., 
  data = na.omit(TitanicSurvival), 
  method = "glm",
  family = "binomial",
  trControl = train_model)
print(modelT)

confusionMatrix(predict(modelT), na.omit(TitanicS$survived))


```

## Example 8 from slides

```{r}

# ProbitModel <- glm(binary.response ∼ Var.1 + Var.2 + ...
# + Var.j, family = binomial(link = "probit"), data =
# data)



```

```{r}

library(ggeffects)

# margins(ProbitModel, variables = "variable.name")

# ggpredict(lm1,terms = "age[0:100 by = 1]") %>% 
#   plot()

```
